---
title: "NOA"
author: "KMicha71"
date: "14 5 2021"
output:
  html_document: 
    keep_md: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Nort Atlantic Oscilation

Download data from:

https://crudata.uea.ac.uk/cru/data/nao/

Jones, P.D., Jónsson, T. and Wheeler, D., 1997: Extension to the North Atlantic Oscillation using early instrumental pressure observations from Gibraltar and South-West Iceland. Int. J. Climatol. 17, 1433-1450. doi: 10.1002/(SICI)1097-0088(19971115)17:13<1433::AID-JOC203>3.0.CO;2-P



```{sh downloads}
 [ -f ./download/nao_3dp.dat ] && mv -f ./download/nao_3dp.dat ./download/nao_3dp.dat.bck
 wget -q -P download https://crudata.uea.ac.uk/cru/data/nao/nao_3dp.dat
 sed -i '1s/^/year,m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11,m12,sum\n/' ./download/nao_3dp.dat
 sed -i 's/^ //g' ./download/nao_3dp.dat 
 sed -i 's/   / /g' ./download/nao_3dp.dat 
 sed -i 's/  / /g' ./download/nao_3dp.dat 
 sed -i 's/ /,/g' ./download/nao_3dp.dat 
 #sed -i '/^"/d' ./download/monthly_in_situ_co2_mlo.csv
 #sed -i '2d;3d' ./download/monthly_in_situ_co2_mlo.csv
```

## Convert to standard format

```{r convert}
library(reshape2)
noa <- read.csv("./download/nao_3dp.dat", sep=",")
noaNew <- melt(noa, id.vars = c("year"))
noaNew <- subset(noaNew, -99 < noaNew$value)
names(noaNew)[names(noaNew) == "variable"] <- "month"
noaNew <- subset(noaNew, 'sum' != noaNew$month)
names(noaNew)[names(noaNew) == "value"] <- "noa1"
noaNew$month <- strtoi(gsub("m", "", noaNew$month))

noaNew$ts <- signif(noaNew$year + (noaNew$month-0.5)/12, digits=6)
noaNew$time <- paste(noaNew$year,noaNew$month, '15 00:00:00', sep='-')

noaNew$noa2 <- round(noaNew$noa1)
noaNew$noa2[noaNew$noa2 < -4] <- -4
noaNew$noa2[noaNew$noa2 > 4] <- 4

noaNew <- noaNew[order(noaNew$time),]

write.table(noaNew, file = "./csv/monthly_noa1.csv", append = FALSE, quote = TRUE, sep = ",",
            eol = "\n", na = "NA", dec = ".", row.names = FALSE,
            col.names = TRUE, qmethod = "escape", fileEncoding = "UTF-8")
```

## Plot diagramm

```{r plot, echo=TRUE}
require("ggplot2")
noa <- read.csv("./csv/monthly_noa1.csv", sep=",")
mp <- ggplot() +
      geom_line(aes(y=noa$noa2, x=noa$ts), color="blue") +
      xlab("Year") + ylab("NOA")
mp

```


https://psl.noaa.gov/gcos_wgsp/Timeseries/Data/rnao.long.data

Luterbacher, J., Xoplaki, E., Dietrich, D., Jones, P.D., Davies, T.D., Portis, D., Gonzalez-Rouco, J.F., von Storch, H., Gyalistras, D., Casty, C., and Wanner, H., 2002. Extending North Atlantic Oscillation Reconstructions Back to 1500. Atmos. Sci. Lett., doi:10.1006/asle.2001.0044.
Luterbacher, J., C. Schmutz, D. Gyalistras, E. Xoplaki, and H. Wanner, 1999: Reconstruction of monthly NAO and EU indices back to AD 1675, Geophys. Res. Lett., 26, 2745-2748.

```{sh downloads2}
 [ -f ./download/rnao.long.data ] && mv -f ./download/rnao.long.data ./download/rnao.long.data.bck
 wget -q -P download https://psl.noaa.gov/gcos_wgsp/Timeseries/Data/rnao.long.data
 
  #remove first line
 sed -i -e 1,1d ./download/rnao.long.data
  sed -i '1s/^/year,m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11,m12\n/' ./download/rnao.long.data
 sed -i 's/^ //g' ./download/rnao.long.data
 sed -i 's/   / /g' ./download/rnao.long.data
 sed -i 's/  / /g' ./download/rnao.long.data
 sed -i 's/ /,/g' ./download/rnao.long.data
 
 cat ./download/rnao.long.data | head -n -4 > ./download/rnao.long.data2
 mv ./download/rnao.long.data2 ./download/rnao.long.data

```

## Convert to standard format

```{r convert2}
library(reshape2)
noa <- read.csv("./download/rnao.long.data", sep=",")
noaNew <- melt(noa, id.vars = c("year"))
noaNew <- subset(noaNew, -99 < noaNew$value)
names(noaNew)[names(noaNew) == "variable"] <- "month"
names(noaNew)[names(noaNew) == "value"] <- "noa1"
noaNew$month <- strtoi(gsub("m", "", noaNew$month))

noaNew$ts <- signif(noaNew$year + (noaNew$month-0.5)/12, digits=6)
noaNew$time <- paste(noaNew$year,noaNew$month, '15 00:00:00', sep='-')

noaNew$noa2 <- round(noaNew$noa1)
noaNew$noa2[noaNew$noa2 < -4] <- -4
noaNew$noa2[noaNew$noa2 > 4] <- 4

noaNew <- noaNew[order(noaNew$time),]

write.table(noaNew, file = "./csv/monthly_noa2.csv", append = FALSE, quote = TRUE, sep = ",",
            eol = "\n", na = "NA", dec = ".", row.names = FALSE,
            col.names = TRUE, qmethod = "escape", fileEncoding = "UTF-8")
```

## Plot diagramm

```{r plot2, echo=TRUE}
require("ggplot2")
noa <- read.csv("./csv/monthly_noa2.csv", sep=",")
mp <- ggplot() +
      geom_line(aes(y=noa$noa2, x=noa$ts), color="blue") +
      xlab("Year") + ylab("NOA")
mp

```


